---
title: "Untitled"
author: "Rebecca Silva"
date: "11/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Desired scenario: 
All parameters except intercept, dose effect, and covariate 3 main effect = 0

* Probability of tox over the 5 doses for patient *with covariate 3 (A)*: $(0.1, 0.3, 0.45, 0.6, 0.8)$

* Probability of tox over the 5 doses for patient *without covariate 3 (B)*: $(0.01, 0.03,0.1,0.15,0.30)$

$$XB =
\begin{bmatrix} 1 & d_1 & 1\\
1 & d_2& 1\\
1 & d_3& 1\\
1 & d_4& 1\\
1 & d_5& 1\\
1 & d_1& 0\\
1 & d_2& 0\\
1 & d_3& 0\\
1 & d_4& 0\\
1 & d_5& 0
\end{bmatrix}*\begin{bmatrix} \beta_0\\
\beta_1\\
\beta_3\\
\end{bmatrix}$$
Set $\beta_2 = \beta_4 = \beta_5 = ...= \beta_{10} = 0$

$$\text{ptox} = 1 - \Phi(X \boldsymbol{\beta})$$
where $\text{ptox}^T =\begin{bmatrix} 0.1, 0.3, 0.45, 0.6, 0.8
0.01, 0.03,0.1,0.15,0.30\\
\end{bmatrix}$

$$\Phi^{-1}(1-\text{ptox}) = X\boldsymbol{\beta}$$

$$\boldsymbol{\beta} = X^{-1} \Phi^{-1}(1-\text{ptox})$$
SCENARIO 1: set (in scenario_seq_true_dose )
SCENARIO 2: X3: mtd 2 and 5 

## SCENARIO 1: 2 and 3

# 1.(a) 2 and 3 previously: 
## w/o X3: 0.079 0.175 *0.321* 0.504 0.685
## w/ X3: 0.162 *0.304* 0.484 0.668 0.818

# 1.(b) 2 and 3: larger jump
## w/o X3: 0.030 0.113 *0.290* 0.543 0.779
## w/ X3: 0.124 *0.311* 0.566 0.796 0.932

# 1.(c) 2 and 3: largest jump: try first 
## w/o X3: 0.012 0.079 *0.290* 0.619 0.877
## w/ X3: 0.089 *0.313* 0.644 0.890 0.981

Code: 
```{r}
true_dose =  c(.1,.3,.5,.7,.9) 
# 1(b) c(0.15, 0.6, 0.6, 0.7, 0.8), c(0.01, 0.01, 0.7, 0.7, 0.8)
# 1(c) c(0.01, 0.8, 0.8, 0.8, 0.95), c(0.01, 0.01, 0.7, 0.7, 0.8)

ptox  = c(c(0.01, 0.8, 0.8, 0.8, 0.95), c(0.01, 0.01, 0.7, 0.7, 0.8))  #c(0.2, 0.4, 0.4, 0.7, 0.8)
y = qnorm(1- ptox)
X = matrix(c(rep(1, 10), rep(c(.1,.3,.5,.7,.9),2), rep(1, 5), rep(0, 5)), nrow= 10)
b = ginv(X) %*% y  # solving for beta params  - rest are 0
# test b 
betas = rep(0, 13)
betas[c(1, 3, 6)] = b
betas

# patient w/ covariate 3 
covars = c(1,0,1,1,0)
ptox_est = rep(0, 5)
for (i in 1:length(true_dose)){
        #set.seed(10)
        ptox_est[i] = 1 - pnorm(betas[1] + betas[3]*true_dose[i] + 
                                    betas[4]*covars[1] + betas[5]*covars[2] + betas[6]*covars[3] + betas[7]*covars[4] +betas[8]*covars[5] 
                            +betas[9]*covars[1]*true_dose[i]+betas[10]*covars[2]*true_dose[i]+betas[11]*covars[3]*true_dose[i] + 
                                    betas[12]*covars[4]*true_dose[i] + betas[13]*covars[5]*true_dose[i])
}
round(ptox_est,3) # 2

# patient w/o covariate 3 
covars = c(1,0,0,1,0)
ptox_est = rep(0, 5)
for (i in 1:length(true_dose)){
        #set.seed(10)
        ptox_est[i] = 1 - pnorm(betas[1] + betas[3]*true_dose[i] + 
                                    betas[4]*covars[1] + betas[5]*covars[2] + betas[6]*covars[3] + betas[7]*covars[4] +betas[8]*covars[5] 
                            +betas[9]*covars[1]*true_dose[i]+betas[10]*covars[2]*true_dose[i]+betas[11]*covars[3]*true_dose[i] + 
                                    betas[12]*covars[4]*true_dose[i] + betas[13]*covars[5]*true_dose[i])
}
round(ptox_est, 3) # 3

```

## SCENARIO 2: 2 and 4
```{r}
true_dose = c(.1,.3,.5,.7,.9)
##c(c(0.15, 0.3, 0.5, 0.6, 0.8), c(0.01, 0.05, 0.12, 0.35, 0.50))
ptox = c(c(0.15, 0.35, 0.5, 0.6, 0.8), c(0.01, 0.05, 0.12, 0.40, 0.50))
y = qnorm(1- ptox)
X = matrix(c(rep(1, 10), rep(c(.1,.3,.5,.7,.9),2), rep(1, 5), rep(0, 5)), nrow= 10)
# first col: for intercept B0
# second col is for B1 which is dose coefficient 
# third col: whether or not patient has covariate 3 (first 5 rows has it, next 5 rows do not)
# a is 10x3 matrix, for 5 different doses * 2 difference possible scenarios (w/ or w/o covarate 3)
# ie here we do not include interactions 


b = ginv(X) %*% y  # solving for beta params  - rest are 0

# test b 
betas = rep(0, 13)
betas[c(1, 3, 6)] = b
betas

# patient w/o covariate 3 
covars = c(1,0,0,1,0)
ptox_est = rep(0, 5)
for (i in 1:length(true_dose)){
        #set.seed(10)
        ptox_est[i] = 1 - pnorm(betas[1] + betas[3]*true_dose[i] + 
                                    betas[4]*covars[1] + betas[5]*covars[2] + betas[6]*covars[3] + betas[7]*covars[4] +betas[8]*covars[5] 
                            +betas[9]*covars[1]*true_dose[i]+betas[10]*covars[2]*true_dose[i]+betas[11]*covars[3]*true_dose[i] + 
                                    betas[12]*covars[4]*true_dose[i] + betas[13]*covars[5]*true_dose[i])
}
round(ptox_est, 3)


# patient w/ covariate 3 
covars = c(1,0,1,1,0)
ptox_est = rep(0, 5)
for (i in 1:length(true_dose)){
        #set.seed(10)
        ptox_est[i] = 1 - pnorm(betas[1] + betas[3]*true_dose[i] + 
                                    betas[4]*covars[1] + betas[5]*covars[2] + betas[6]*covars[3] + betas[7]*covars[4] +betas[8]*covars[5] 
                            +betas[9]*covars[1]*true_dose[i]+betas[10]*covars[2]*true_dose[i]+betas[11]*covars[3]*true_dose[i] + 
                                    betas[12]*covars[4]*true_dose[i] + betas[13]*covars[5]*true_dose[i])
}
round(ptox_est,3)
```

## SCENARIO 3: 2 and 5
```{r}
true_dose = c(.1,.3,.5,.7,.9)
# c(c(0.15, 0.45, 0.5, 0.6, 0.8), c(0.01, 0.05, 0.1, 0.12, 0.45))
ptox = c(c(0.15, 0.3, 0.5, 0.6, 0.8), c(0.01, 0.05, 0.1, 0.12, 0.30))
y = qnorm(1- ptox)
X = matrix(c(rep(1, 10), rep(c(.1,.3,.5,.7,.9),2), rep(1, 5), rep(0, 5)), nrow= 10)
# first col: for intercept B0
# second col is for B1 which is dose coefficient 
# third col: whether or not patient has covariate 3 (first 5 rows has it, next 5 rows do not)
# a is 10x3 matrix, for 5 different doses * 2 difference possible scenarios (w/ or w/o covarate 3)
# ie here we do not include interactions 


b = ginv(X) %*% y  # solving for beta params  - rest are 0

# test b 
betas = rep(0, 13)
betas[c(1, 3, 6)] = b
betas

# patient w/o covariate 3 
covars = c(1,0,0,1,0)
ptox_est = rep(0, 5)
for (i in 1:length(true_dose)){
        #set.seed(10)
        ptox_est[i] = 1 - pnorm(betas[1] + betas[3]*true_dose[i] + 
                                    betas[4]*covars[1] + betas[5]*covars[2] + betas[6]*covars[3] + betas[7]*covars[4] +betas[8]*covars[5] 
                            +betas[9]*covars[1]*true_dose[i]+betas[10]*covars[2]*true_dose[i]+betas[11]*covars[3]*true_dose[i] + 
                                    betas[12]*covars[4]*true_dose[i] + betas[13]*covars[5]*true_dose[i])
}
round(ptox_est,3)


# patient w/ covariate 3 
covars = c(1,0,1,1,0)
ptox_est = rep(0, 5)
for (i in 1:length(true_dose)){
        #set.seed(10)
        ptox_est[i] = 1 - pnorm(betas[1] + betas[3]*true_dose[i] + 
                                    betas[4]*covars[1] + betas[5]*covars[2] + betas[6]*covars[3] + betas[7]*covars[4] +betas[8]*covars[5] 
                            +betas[9]*covars[1]*true_dose[i]+betas[10]*covars[2]*true_dose[i]+betas[11]*covars[3]*true_dose[i] + 
                                    betas[12]*covars[4]*true_dose[i] + betas[13]*covars[5]*true_dose[i])
}
round(ptox_est,3)
```

## SCENARIO 4: No covar


# Scenario 5: 2 covars, far:
## (0,0), (1,1): 0.01, 0.05, 0.1, 0.15, *0.3*
## (1,0), (0,1): 0.10, *0.30*, 0.50, 0.70, 0.90

# Scenario 6: 2 covars, close:
## (0,0), (1,1): 0.012 0.079 *0.290* 0.619 0.877
## (1,0), (0,1): 0.089 *0.313* 0.644 0.890 0.981


# SCENARIO 7: 1 adn 3 checking boundaries 
```{r}
ptox  = c(c(0.4, 0.7, 0.7, 0.8, 0.9), c(0.01, 0.10, 0.35, 0.6, 0.8))
y = qnorm(1- ptox)
X = matrix(c(rep(1, 10), rep(c(.1,.3,.5,.7,.9),2), rep(1, 5), rep(0, 5)), nrow= 10)
```
## w/o X3: 0.044 0.127 *0.281* 0.493 0.708
## w/ X3: *0.294* 0.508 0.720 0.874 0.956

# SCENARIO 8: 2 covars
in gen_2cov_fn.R
2 groups: (0,0) mtd = 2 rest mtd = 5
3 groups: (0,0) mtd = 2 (1,0) and (0,1) mtd = 4, (1,1) mtd = 5




Others: 
# one covar matters but not sparse (one other covar is not 0)
```{r}
# cov 1 and 3: c(.1, .3, .45, .6,.8) # mtd 3
# cov 1 not 3: c(0.01, .1, .3, .45, .6) # mtd 4 
# cov 3 not 1: c(.1, .3, .45, .6, .8) # mtd 3
# not 3 nor 1: c(.01, .07, .1, .3, .45) # mtd 4 
# only main effects: no interaction for cov 3 nor 1


true_dose = c(.1,.3,.5,.7,.9)
ptox = c(c(.05, .05, .15, .6,.8), c(0.01, .1, .1, .5, .6), c(.1, .15, .3, .6, .8), c(0.01, .1, .1, .2, .3))
y = qnorm(1- ptox)
X = matrix(c(rep(1, 20), rep(c(.1,.3,.5,.7,.9),4), rep(1, 5), rep(1, 5),rep(0, 5), rep(0, 5), rep(1, 5), rep(0, 5),rep(1, 5), rep(0, 5)),  nrow= 20)



b = ginv(X) %*% y  # solving for beta params  - rest are 0

# test b 
betas = rep(0, 13)
betas[c(1, 3, 4, 6)] = b # 1 = intercept, 3 = dose effect, 4 = cov 1, 6 = cov 3
betas

# patient w/ covariate 3 and 1
covars = c(1,0,1,1,0)
ptox_est = rep(0, 5)
for (i in 1:length(true_dose)){
        #set.seed(10)
        mat_B0 = matrix(c(1,0,true_dose[i]), ncol = 3) # intercept, sd, true
        mat_covars = matrix(covars, ncol = length(covars))
        matX = matrix(cbind(mat_B0, mat_covars, mat_covars*true_dose[i]), ncol = 1) # mat of int, dose, cov, cov*dose
        matB = matrix(betas, ncol = length(betas))
        ptox_est[i] = 1 - pnorm(matB %*% matX)
}
ptox_est

# patient w/o covariate 3, with 1 
covars = c(1,0,0,1,0)
ptox_est = rep(0, 5)
for (i in 1:length(true_dose)){
        #set.seed(10)
        mat_B0 = matrix(c(1,0,true_dose[i]), ncol = 3) # intercept, sd, true
        mat_covars = matrix(covars, ncol = length(covars))
        matX = matrix(cbind(mat_B0, mat_covars, mat_covars*true_dose[i]), ncol = 1) # mat of int, dose, cov, cov*dose
        matB = matrix(betas, ncol = length(betas))
        ptox_est[i] = 1 - pnorm(matB %*% matX)
}
ptox_est

# patient w/ covariate 3 not 1
covars = c(0,0,1,1,0)
ptox_est = rep(0, 5)
for (i in 1:length(true_dose)){
        #set.seed(10)
        mat_B0 = matrix(c(1,0,true_dose[i]), ncol = 3) # intercept, sd, true
        mat_covars = matrix(covars, ncol = length(covars))
        matX = matrix(cbind(mat_B0, mat_covars, mat_covars*true_dose[i]), ncol = 1) # mat of int, dose, cov, cov*dose
        matB = matrix(betas, ncol = length(betas))
        ptox_est[i] = 1 - pnorm(matB %*% matX)
}
ptox_est # wrong mtd 

# patient w/o 3 w/o 1
covars = c(0,0,0,1,0)
ptox_est = rep(0, 5)
for (i in 1:length(true_dose)){
        #set.seed(10)
        mat_B0 = matrix(c(1,0,true_dose[i]), ncol = 3) # intercept, sd, true
        mat_covars = matrix(covars, ncol = length(covars))
        matX = matrix(cbind(mat_B0, mat_covars, mat_covars*true_dose[i]), ncol = 1) # mat of int, dose, cov, cov*dose
        matB = matrix(betas, ncol = length(betas))
        ptox_est[i] = 1 - pnorm(matB %*% matX)
}
ptox_est
```

Interactions?

